---
title: "Field Notes"
---

```{r}
#| include: false
library(tidyverse)
library(plotly)
parks <- read_csv("data/park_weather_monthly.csv", show_col_types = FALSE)
```

::: {.hero}
<div class="hero-kicker">A decade in the wild · 2007—2016</div>

# When the forecast becomes part of the itinerary

Thousands of monthly visits to 60 US national parks, set beside the temperatures and precipitation that shape a day outside. Explore the broad pattern, then step inside one park at a time.

::: {.metric-row}
<div class="metric"><strong>`r scales::comma(nrow(parks))`</strong><span>park-month observations</span></div>
<div class="metric"><strong>`r n_distinct(parks$park_name)`</strong><span>national parks</span></div>
<div class="metric"><strong>`r min(parks$year)`–`r max(parks$year)`</strong><span>years covered</span></div>
:::
:::

## The national picture

Each point is one park-month. The warm band is busy: visitation rises as outdoor conditions become more comfortable, but the relationship is not a straight line. Desert parks often trade summer heat for winter crowds, while high-latitude parks compress their season into a few bright months.

```{r}
#| label: fig-overview
#| fig-cap: "Monthly visitors and average temperature across the park network. Point size reflects precipitation."
overview <- parks |>
  mutate(year = factor(year))

plot_ly(
  overview,
  x = ~average_temperature_f,
  y = ~visitor_count,
  type = "scatter",
  mode = "markers",
  color = ~year,
  colors = "YlOrRd",
  size = ~precipitation_in,
  sizes = c(6, 28),
  opacity = .55,
  text = ~paste(park_name, month_name, year, sep = " · "),
  hovertemplate = "%{text}<br>%{x:.0f}°F · %{y:,.0f} visits<extra></extra>"
) |>
  layout(
    xaxis = list(title = "Average monthly temperature (°F)", zeroline = FALSE),
    yaxis = list(title = "Monthly visitor count", tickformat = ","),
    legend = list(title = list(text = "Year"))
  )
```

::: {.callout}
**Read the shape, not just the average.** Weather helps set the stage, but access, wildfire, road closures, school calendars, and park size also determine who can show up.
:::

## Start exploring

Use the [Exploration](exploration.qmd) tab to select an individual park, zoom into a time period, and compare its visitor curve with local temperature and precipitation. The [Story](story.qmd) follows one focused question: why do northern and desert parks peak in opposite seasons?

<p class="source-note">Data note: annual visitation totals come from the National Park Service Stats archive via TidyTuesday. Monthly counts and weather values are a reproducible teaching approximation derived from annual totals and regional climate profiles; see the data-preparation script for the method.</p>