---
title: "Story"
---

```{r}
#| include: false
library(tidyverse)
library(plotly)
parks <- read_csv("data/park_weather_monthly.csv", show_col_types = FALSE)
```

<div class="hero">
<div class="hero-kicker">A seasonal field note</div>

# The sweet spot moves with the map

There is no single national park season. In the north, winter closes the door and summer opens it. In the desert, summer heat does the opposite: the comfortable months arrive after the crowds have gone.
</div>

## 01 / Summer is not one thing

```{r}
#| label: fig-season-contrast
#| fig-cap: "Estimated summer and winter visitor totals for two northern and two desert parks."
story_parks <- c("Glacier", "Acadia", "Death Valley", "Zion")
seasonal <- parks |>
  filter(park_name %in% story_parks) |>
  mutate(season = if_else(month %in% 6:8, "Summer", "Winter")) |>
  group_by(park_name, season) |>
  summarise(visitors = sum(visitor_count), .groups = "drop") |>
  mutate(park_name = forcats::fct_reorder(park_name, visitors, .fun = sum))

ggplot(seasonal, aes(park_name, visitors, fill = season)) +
  geom_col(position = "dodge", width = .7) +
  scale_fill_manual(values = c("Summer" = "#d6674d", "Winter" = "#8da99a")) +
  scale_y_continuous(labels = scales::label_number(scale = 1e-6, suffix = "M")) +
  labs(x = NULL, y = "Visitors (millions)", fill = NULL) +
  theme_minimal(base_family = "Manrope") +
  theme(panel.grid.minor = element_blank(), legend.position = "top")
```

Glacier and Acadia concentrate their welcome into warm months. Zion and Death Valley spread their opportunity differently: heat changes the timing of a visit, rather than erasing demand altogether. A national average would flatten this useful contrast.

## 02 / Peaks are weather-shaped

```{r}
#| label: fig-weather-shifts
#| fig-cap: "Visitor counts and temperature profiles for Glacier and Death Valley."
events <- parks |>
  filter(park_name %in% c("Glacier", "Death Valley")) |>
  group_by(park_name, month_name, month) |>
  summarise(visitors = sum(visitor_count), temperature = mean(average_temperature_f), .groups = "drop")

ggplot(events, aes(month, visitors, color = park_name)) +
  geom_line(linewidth = 1.2) + geom_point(size = 2) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  scale_y_continuous(labels = scales::label_number(scale = 1e-6, suffix = "M")) +
  scale_color_manual(values = c("Glacier" = "#d6674d", "Death Valley" = "#e7a24b")) +
  labs(x = NULL, y = "Monthly visitors (millions)", color = NULL) +
  theme_minimal(base_family = "Manrope") +
  theme(panel.grid.minor = element_blank(), legend.position = "top")
```

The curves are mirror images because the operational seasons are mirror images. Snowpack, road access, heat exposure, and daylight all turn a temperature reading into a practical decision: when can visitors arrive, move through the landscape, and stay comfortable long enough to explore?

## A picture of the shoulder season

::: {.columns}
::: {.column width="50%"}
![A mountain landscape in a cool-season light.](https://images.unsplash.com/photo-1500534623283-312aade485b7?auto=format&fit=crop&w=1200&q=80)
:::
::: {.column width="50%"}
![A warm desert landscape under open sky.](https://images.unsplash.com/photo-1474044159687-1ee9f3a51722?auto=format&fit=crop&w=1200&q=80)
:::
:::

<p class="source-note">Interpretation: these visuals use modeled monthly weather profiles, not event-level NOAA observations. They are best read as a clear demonstration of the seasonal mechanism, not as a causal estimate of weather's independent effect.</p>